---
title: "Dockstore and Bioconductor for AnVIL"
author: "BJ Stubbs, S Gopaulakrishnan, Nitesh Turaga, Martin Morgan, Vincent J. Carey, stvjc at channing.harvard.edu"
date: "`r format(Sys.time(), '%B %d, %Y')`"
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Dockstore and Bioconductor}
  %\VignetteEncoding{UTF-8}
output:
  BiocStyle::html_document:
    highlight: pygments
    number_sections: yes
    theme: united
    toc: yes
  BiocStyle::pdf_document:
    toc: yes
    number_sections: yes
---

# Introduction: Basic concepts of Dockstore and Bioconductor

```{r setup,echo=FALSE}
suppressPackageStartupMessages({
library(AnVIL) # get from https://github.com/Bioconductor/AnVIL
library(dplyr)
})
```

Dockstore is the "VM/Docker sharing infrastructure and management component"
of the Global Alliance for Genomics and Health (GA4GH).
Dockstore.org implements the
infrastructure by defining APIs for coupling Docker images
with formalized workflow specifications.
The application
of this concept to the PanCancer Analysis of Whole Genomes
(PCAWG) is described in a
[2017 paper](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5333608/) by
O'Connor and colleagues.

[Bioconductor](http://www.bioconductor.org)
is a software ecosystem based in the R language
for the analysis and comprehension of genome-scale experiments.
[An overview](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4509590/)
was published in 2015.

Bioconductor implemented a ["common workflow"](https://www.bioconductor.org/packages/release/BiocViews.html#___Workflow) concept a number
of years ago.  (The term "common" is used on the Bioconductor
landing page, to indicate that many bioinformaticians would
be expected to engage with tasks reviewed in these workflows.
The "common workflow" phrase is not intended to indicate
a relationship to CWL, the "common workflow language" standard.)

In Bioconductor parlance, a "common workflow" is an R package
accompanied by a markdown or Sweave vignette that narrates
and executes the workflow tasks using R.  Any sufficiently
endowed deployment of R on a sufficient computing platform
will run the workflow tasks to completion; this assertion is
checked on a regular basis through Bioconductor's continuous
integration system.  When the assertion fails, the workflow
maintainer is alerted and corrections are made.

Advantages to marrying the Bioconductor workflow concept with
Dockstore VM/Docker/workflow infrastructure include

- reduction in user burden of configuring and maintaining the
execution platform
- utilization of parameterized formal workflow specification in CWL,
WDL, or Nextflow
- general advantages to users of fostering participation in GA4GH
best practices related to reproducibility and transparency

Open questions concern the balance between specification of
workflow steps in R and in the formal workflow language.  Bioconductor
workflows can be written to take advantage of R's capabilities
to drive computations on potentially heterogeneous
clusters with programmable fault tolerance and job control.
The particular advantages of CWL/WDL/Nextflow and
other aspects of the Dockstore ecosystem need to be experienced,
measured, and documented to help developers establish the
appropriate balance between programming R and programming
an exogenous workflow environment.

# Working with the Dockstore API in Bioconductor

The [AnVIL](https://github.com/Bioconductor/AnVIL) package
handles basic aspects of authentication and API element
cataloguing for the AnVIL project.

```{r lka}
library(AnVIL)
```

An object called 'dockstore' is defined and will be
used to process API requests.  Groups of API components are
obtained via `tags()`.

```{r getmeths}
knitr::kable(tags(dockstore) %>% distinct(tag))
```
We're interested in
the 'users' component.  Higher level methods will be
introduced to help here, but for now we stick to base R
methods.

```{r lku}
tags(dockstore) %>% filter(tag == "users") %>%
    pull(operation) %>% sort()
```

We can use the following to determine our user identifier.
```{r lklk}
myuid = (dockstore$getUser() %>% as.list)$id
myuid
```
